Load packages
```{r}
library(tidyverse)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(skimr)
library(naniar)
```

Read in data
```{r}
grades = read_csv("class-grades.csv")
```

Structure and summary
```{r}
str(grades)
summary(grades)
skim(grades)
```
Looking at missingness by variable and combinations of missingness using "aggr" from VIM package.  
```{r}
vim_plot = aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
#the cex.axis reduces size of text on x-axis so labels fit better
```

A view of missingness by variable and row.  
```{r}
vis_miss(grades) #from the naniar package
```

Row-wise deletion of any row with at least one NA:
```{r}
grades_rowdel = grades %>% drop_na() 
#alternatively can specify which variable(s) on which to do the drop_na
#drop_na(Age)
skim(grades_rowdel)
```

Column-wise deletion  
```{r}
grades_coldel = grades %>% select(-Final, -TakeHome, -Midterm, -Tutorial) 
vim_plot = aggr(grades_coldel, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
skim(grades_coldel)
```

Using "mice" package -- m=5 and method=pmm and random number seed 123
```{r}
set.seed(123) #sets seed for random number generator
imp_grades = mice(grades, m=5, method='pmm', printFlag=FALSE)
#m is the number of imputations, 5 is a reasonable value as a default
#pmm is "predictive mean matching" = imputation method for numeric data
#printFlag reduces amount of output
summary(imp_grades)
```

Merge the imputed values into grades data frame  
```{r}
grades_complete = complete(imp_grades) 
summary(grades_complete)
```