Needed libraries  
```{r}
library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)
```

Read in data and clean
```{r}
parole = read_csv("parole.csv")

parole = parole %>% mutate(male = as_factor(male)) %>% 
  mutate(male = fct_recode(male, "female" = "0", "male" = "1" )) 

parole = parole %>% mutate(race = as_factor(race)) %>% 
  mutate(race = fct_recode(race, "white" = "1", "other" = "2" )) 


parole = parole %>% mutate(state = as_factor(state)) %>% 
  mutate(state = fct_recode(state, "Kentucky" = "2", "Louisiana" = "3", "Virginia" = "4", "other" = "1" )) 


parole = parole %>% mutate(crime = as_factor(crime)) %>% 
  mutate(crime = fct_recode(crime, "larceny" = "2", "drug-related" = "3", "driving-related" = "4", "other" = "1" )) 


parole = parole %>% mutate(multiple.offenses = as_factor(multiple.offenses)) %>% 
  mutate(multiple.offenses = fct_recode(multiple.offenses, "No" = "0", "Yes" = "1" )) 


parole = parole %>% mutate(violator = as_factor(violator)) %>% 
  mutate(violator = fct_recode(violator, "No" = "0", "Yes" = "1" )) 
  
```


How many of the parolees ended up violating parole?
```{r}
nrow(parole[parole$violator == 'Yes', ])
```


Split the data (training and testing). 70% of the data to training. Stratified the random split by the response variable "violator". Random number seed set is 1234.
```{r}
set.seed(12345)
parole_split = initial_split(parole, prop = 0.70, strata = violator)
train = training(parole_split)
test = testing(parole_split)
```

Check levels
```{r}
levels(train$violator)
```

Rearrange factor levels to that the positive class is listed second. Not necessary since we got 0 1 (1 is positive) but helpful to remember
```{r}
train = train %>% mutate(violator = fct_relevel(violator, c("No","Yes")))
levels(train$violator)
```

Voilator by Gender  
```{r}
ggplot(train, aes(x=male, fill = violator)) + geom_bar() + theme_bw() +
  facet_grid(. ~ violator)
```

Violator by State  
```{r}
ggplot(train, aes(x= state, fill = violator)) + geom_bar() + theme_bw() +
  facet_grid(. ~ violator)
```

Voilator by Max Sentence
```{r}
ggplot(train, aes(x=max.sentence, fill=violator)) + geom_bar() +
  facet_grid(. ~ violator)
```

Building a logistic regression model with state to predict violator
```{r}
parole_model = 
  logistic_reg() %>% #logistic reg
  set_engine("glm") #standard logistic regression engine

parole_recipe = recipe(violator ~ state, train) %>%
  step_dummy(all_nominal(), -all_outcomes()) #exclude the response variable from being dummy converted  

logreg_wf = workflow() %>%
  add_recipe(parole_recipe) %>% 
  add_model(parole_model)

parole_fit = fit(logreg_wf, train)
```

```{r}
summary(parole_fit$fit$fit$fit)
```

New recipe using state, multiple.offenses and race to predict violator
```{r}
parole_recipe2 = recipe(violator ~ state + multiple.offenses + race, train) %>%
  step_dummy(all_nominal(), -all_outcomes()) #exclude the response variable from being dummy converted  

logreg_wf2 = workflow() %>%
  add_recipe(parole_recipe2) %>% 
  add_model(parole_model)

parole_fit2 = fit(logreg_wf2, train)
```

```{r}
summary(parole_fit2$fit$fit$fit)
```


Using logreg_wf2 -- predict the probability that the a parolee from Louisiana, that has multiple offences, and is white, will violate parole.
```{r}
newdata = data.frame(state = "Louisiana", multiple.offenses = "Yes", race = "white")
predict(parole_fit2, newdata, type="prob")
```


ROC curve and probability threshold that best balances specificity and sensitivity for the training set
```{r}
predictions = predict(parole_fit2, train, type="prob") #develop predicted probabilities
head(predictions)
```

Extract just the "Yes" prediction.  
```{r}
predictions = predict(parole_fit2, train, type="prob")[2]
head(predictions)
```

Threshold selection  
```{r}
#Change this next line to the names of your predictions and the response variable in the training data frame
ROCRpred = prediction(predictions, train$violator) 

###You shouldn't need to ever change the next two lines:
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```

Area under the curve (AUC). AUC is a measure of the strength of the model. Values closer to 1 are better. Can be used to compare models.  
```{r}
as.numeric(performance(ROCRpred, "auc")@y.values)
```

```{r}
#Determine threshold to balance sensitivity and specificity
#DO NOT modify this code
opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperf, ROCRpred))
```

Test thresholds to evaluate accuracy  
```{r}
#confusion matrix
#The "No" and "Yes" represent the actual values
#The "FALSE" and "TRUE" represent our predicted values
t1 = table(train$violator,predictions > 0.2015788)
t1
```

Sensitivity
```{r}
36/(18+36)
```


Calculate accuracy  
```{r}
(t1[1,1]+t1[2,2])/nrow(train)
```

Testing accuracy for different thresholds

Threshold = 0.2  
```{r}
t1 = table(train$violator,predictions > 0.3)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

Threshold = 0.3
```{r}
t1 = table(train$violator,predictions > 0.3)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

Threshold = 0.4  
```{r}
t1 = table(train$violator,predictions > 0.4)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

Threshold = 0.5  
```{r}
t1 = table(train$violator,predictions > 0.5)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

